### 2019-10-08 Progress Documentation.Rmd
added progress report

---
title: "Progress Documentation"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{itemize}
  \item 10 / 08 / 2019: Nimble crashes my Mac laptop every now and then, so I should do more simulation with that.
\end{itemize}