### 2020-04-27 PPP/Figures.Rmd
ppp update and adding accuracy bench

---
title: "Figures"
author: "Evan Miyakawa"
output: pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

```

```{r}
library(tidyverse); options(tibble.width = Inf)
library(magrittr)
library("greta")
```

```{r}
set.seed(2)

df <- tibble(x = rnorm(750, mean = 50, sd = 25), y = rnorm(750, mean = x * 1.2, sd = 5))

df %>% ggplot(aes(x,y)) +
  geom_point(alpha = 0.7) + 
  theme_minimal() + 
  theme(
    axis.title = element_blank(),
    axis.text = element_blank()
  )
```

```{r}
set.seed(2)

df <- tibble(x = rnorm(750, mean = 50, sd = 25), y = rnorm(750, mean = x * 1.2, sd = 5))

df %>% ggplot(aes(x,y)) +
  geom_point(alpha = 0.3) + 
  theme_minimal() + 
  theme(
    axis.title = element_blank(),
    axis.text = element_blank()
  )
```

```{r, fig.height = 3}

log_df <- tibble(x = seq(-4,4,length.out = 100), d_norm = - dnorm(x))

log_df %>% ggplot(aes(x,d_norm)) + 
  geom_line(size = 2) + 
  theme_minimal() + 
  theme(
    axis.title = element_blank(),
    axis.text = element_blank()
  )

```

```{r}

## generate/specify data
################################################################################

alpha <- 2 #beta distribution alpha
beta <- 1 #beta distribution beta
N <- 10 # sample size

set.seed(1)

n <- rbinom(N, 20, 0.5)
theta <- rbeta(N, alpha, beta)
y <- rbinom(N, n, theta)

n <- as_data(n)
y <- as_data(y)

## specify greta model
################################################################################

theta <- beta(2,1, dim = c(N,1))
distribution(y) <- binomial(n, theta, dim = c(N,1))

greta_model <- model(theta)

plot(greta_model)

```

```{r, eval = FALSE}
p <- .25 # binomial p 
n <- 10 # binomial n

set.seed(1)

(y <- rbinom(1, n, p))

model_data <- list(
  "y" = y,
  "N" = n 
)
```

```{r, eval = FALSE}
n_chains <- 4L
n_iter <- 1e4L
n_warmup <- 1e3L
```

```{r, eval = FALSE}
## load library
library(R2OpenBUGS)

## model script code
bugs_model <- function() {
  y ~ dbin(p,N)
  p ~ dbeta(1,1)
}

## parameters to monitor
bugs_monitor <- "p"

## temporary model script text file creation
bugs.file <- file.path(tempdir(), "model.txt")
write.model(bugs_model, bugs.file)

## run the model
bugs_fit <- bugs(
    "model.file" = bugs.file, "data" = model_data, "parameters.to.save" = bugs_monitor, 
    "inits" = NULL, "n.chains" = n_chains, "n.iter" = n_iter, "n.burnin" = n_warmup
)
```

```{r, eval = FALSE}
## load libraries
library("rjags"); library("runjags")

## model script code
jags_model <- "
  model{
    y ~ dbin(p,n)
    p ~ dbeta(1,1)
  }
"

## parameters to monitor
jags_monitor <- c("p")

## run the model
jags_fit <- run.jags(
    "model" = jags_model, "data" = model_data, "monitor" = jags_monitor, 
    "n.chains" = n_chains, "sample" = n_iter, "burnin" = n_warmup
) 
```

```{r, eval = FALSE}
## load library
library(nimble)

## model script code
nimble_model <- nimbleCode({
  y ~ dbin(p,n)
  p ~ dbeta(1,1)
})

## parameters to monitor
nimble_monitor <- c("p")

## specify initial values of parameters
nimble_inits <- list(
  "p" = rbeta(1,1,1)
)

## run the model
nimble_fit <- nimbleMCMC(
    "code" = nimble_model, "data" = model_data, 
    "inits" = nimble_inits, "monitors" = nimble_monitor, "nchains" = n_chains, 
    "niter" = n_iter, "nburnin" = n_warmup, "summary" = TRUE
)
```

```{r, eval = FALSE}
## .stan file

data {
  int<lower=0> y;
  int<lower=0> n;
}

parameters {
  real p;                
}

model {
  y ~ binomial(n,p);
  p ~ beta(1,1);
}
```

```{r, eval = FALSE}
## load library
library("rstan")

## run the model
stan_fit <- stan(
    "file" = "stan_file.stan", "data" = model_data, 
    "chains" = n_chains, "iter" = n_iter, "warmup" = n_warmup
)

```

```{r, eval = FALSE}
## convert data into greta arrays
y <- greta::as_data(y)
n <- greta::as_data(n)
```

```{r}
## load library
library(greta)

## specify prior for parameter(s) of interest
p <- beta(1,1)

## define distribution of data variable and relationship with parameter
distribution(y) <- binomial(n,p)

## compile model and run
greta_model <- model(p)

```

```{r, eval = FALSE}
greta_fit <- mcmc(
    "model" = greta_model, "n_samples" = n_iter,
    "warmup" = n_warmup,"chains" = n_chains
)

```