### 2020-04-30 Simulation Results/Simulation Results.Rmd
add large version of hierch full beta binom

---
title: "Simulation Results"
author: "Evan Miyakawa"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
library(tidyverse); options(tibble.width = Inf)
library(magrittr)
library(here)
```

```{r}
exp_gamma_bench_files <- c(
  here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-results.rds"),
  here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-results-greta.rds"),
  here("Simulation Results", "simple-models", "continuous", "exponential", "gamma-prior", "exponential-gamma-results-stan.rds")
)

normal_gamma_tau_bench_files <- c(
  here("Simulation Results", "simple-models", "continuous", "normal", "gamma-prior-tau", "normal-gamma-tau-results.rds"),
  here("Simulation Results", "simple-models", "continuous", "normal", "gamma-prior-tau", "normal-gamma-tau-results-greta.rds"),
  here("Simulation Results", "simple-models", "continuous", "normal", "gamma-prior-tau", "normal-gamma-tau-results-stan.rds")
)

normal_normal_bench_files <- c(
  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-results.rds"),
  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-results-greta.rds"),
  here("Simulation Results", "simple-models", "continuous", "normal", "normal-prior", "normal-normal-results-stan.rds")
)

beta_bernoulli_bench_files <- c(
  here("Simulation Results", "simple-models", "discrete", "bernoulli", "beta-prior", "beta-bernoulli-results.rds"),
  here("Simulation Results", "simple-models", "discrete", "bernoulli", "beta-prior", "beta-bernoulli-results-greta.rds"),
  here("Simulation Results", "simple-models", "discrete", "bernoulli", "beta-prior", "beta-bernoulli-results-stan.rds")
)

beta_binomial_bench_files <- c(
  here("Simulation Results", "simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-results.rds"),
  here("Simulation Results", "simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-results-greta.rds"),
  here("Simulation Results", "simple-models", "discrete", "binomial", "beta-prior", "beta-binomial-results-stan.rds")
)

poisson_gamma_bench_files <- c(
  here("Simulation Results", "simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-results.rds"),
  here("Simulation Results", "simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-results-greta.rds"),
  here("Simulation Results", "simple-models", "discrete", "poisson", "gamma-prior", "poisson-gamma-results-stan.rds")
)

combine_bench <- function(files, model_name) {
  combined_results <- files %>% map_dfr(~ readRDS(.))
  combined_results

  expression_string <- combined_results$expression[[7]]
  expression_string
  get_method_name <- function(expression_string) {
    string <- as.character(expression_string)[1]
    string
    if (string == "run.jags") {
      method <- "JAGS"
    } else if (string == "bugs") {
      method <- "OpenBUGS"
    } else if (string == "nimbleMCMC") {
      method <- "Nimble"
    } else if (string == "stan") {
      method <- "STAN"
    } else if (string == "mcmc") {
      method <- "Greta"
    } else if (string == "stan_model") {
      method <- "STAN Compile"
    }
    method
  }

  method_names <- combined_results$expression %>% map_chr(get_method_name)

  method_mean <- combined_results$time %>% map_dbl(~ mean(.))
  method_sd <- combined_results$time %>% map_dbl(~ sd(.))

  results <- combined_results %>% 
    mutate(method = method_names, mean = method_mean, sd = method_sd, 
           model_name = model_name) %>% 
    select(method, mean, median, min, sd, n_itr, model_name)

  rows_to_keep <- results$method %>% unique() %>% 
    map_dbl(~ which(results$method == .)[1])

  results <- results[rows_to_keep,] %>% 
    arrange(method)

}

exp_gamma_results <- exp_gamma_bench_files %>% 
  combine_bench("exponential_gamma")
normal_gamma_tau_results <- normal_gamma_tau_bench_files %>%
  combine_bench("normal_gamma_tau")
normal_normal_results <- normal_normal_bench_files %>%
  combine_bench("normal_normal")
beta_bernoulli_results <- beta_bernoulli_bench_files %>%
  combine_bench("beta_bernoulli")
beta_binomial_results <- beta_binomial_bench_files %>%
  combine_bench("beta_binomial")
poisson_gamma_results <- poisson_gamma_bench_files %>%
  combine_bench("poisson_gamma")
poisson_gamma_results

x <- readRDS(bench_files[3])
x
```